Paper Title*\\ {\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and should not be used} \thanks{Identify applicable funding agency here. If none, delete this.}

### 2019-05-13 icdm/submission/includes/discussion.tex
discussion finished reworked

\subsubsection{Large differences with NAG and SGD when using shuffling}
SGD and NAG both have very different values when using shuffling in comparison to unshuffled input, as shown in figures \ref{fig:results_sgd} and \ref{fig:results_nag}, plain SGD. We do not only attribute those results to input shuffling but partially also to a good random initialization guess. Due to a lack of time these results were not replicate more than once.

\subsection{Comparison to Gensim}
This Section will compare our finding extensively to Gensim \cite{gensim}. As explained in Section \ref{ssec:gensim}, Gensim is optimized to have a very high throughput, this made it possible to achieve a lot of computations. Furthermore, Gensim provides access to the loss and the resulting word embeddings, which facilitated the comparison process.
This Section is Structured as follows:First it will describe the used configuration of Gensim, second it will compare Gensim to our implementation of the SGNS with the use of SGD and finally compare our model with Adam to Gensim. A graphic showing the comparison of our models to Gensim can be found in Figure \ref{fig:gensim_vs_adam}.

\iffalse
As Mikolov et al. published the original paper which introduced the SGNS, it is, of course, relevant to compare ourselves to their work. The first important thing to take under account is that Mikolov et al. only trained their model on a very large google news dataset incorporating more than 3 billion words. This makes the comparison of our work more difficult. But we will make some assumptions, as they could be of value if true.
In their original paper, Mikolov et al. reported results from computations that took 1 and 3 epochs. We accord these good results, which clearly outperformed our SGD model and Gensim, to the very large dataset and furthermore, as a matter of fact, their results are better with 3 than with 1 epoch. We do not have any information about the convergence time or criterion. Hence it would be interesting to use their dataset for comparison. \\
One thing we can compare is the quality of our word embeddings. Mikolov et al. did not report any results on their model with the text8 dataset, but they, therefore, published their code. Which was then used by Ji et al \cite{intel} and they tested the model on the text8 dataset. They reported a word similarity of 0.63 on the word similarity task. This is obviously outperformed by all our models. We did not find any explanation on why those results differ as much. \\
The final assumption is that an advanced optimizer could maybe outperform SGD in terms of quality on a large data set. This will be discussed in further work.

\fi

\subsubsection{Configuration of Gensim}
To compare our-self in a correct  manner we used the same dataset, with the same preprocessing parameters, i.e subsampling and min count. The hyper parameters of the network (window size, embedding size, number of negative sample, exponent to which the unigram distribution is raised to chose neg samples) are equivalent to ours, described in Section \ref{ssec:config}. The only difference with our model is the learning rate. Gensim has a starting learning rate of $0.025$ and linearly decrease it at every epoch to $0.0001$.

\subsubsection{Gensim vs. SGD}
As stated earlier, we are not going to compare this work to Gensim in run time. Gensim is heavily optimized and written in cython\footnote{https://rare-technologies.com/word2vec-in-python-part-two-optimizing/}, which is 23x faster than plain Numpy. Since we used PyTorch the difference is not that big, but still remains. As shown in Figure \ref{fig:gensim_vs_adam} the convergence time was not the same between our implementation and Gensim. There are different possible reasons why this could be the case:\\ First, our batched approach could hinder performance in terms of convergence since our loss function is not exactly the same. When a word appears more than once in our batch, the gradient will be an average over the gradients of each pair alone, as it is done by Gensim.\\ Secondly, a difference to our implementation is the fact that Gensim checks whether negative samples are not equal to the context word. If that is the case Gensim selects a new random sample. Therefore, the learning of the input and output context is optimized. \\Finally, another possibility is the decay of the learning rate used by Gensim. In fact, decaying the learning rate has been proven in a lot of work to decrease the convergence time. Gensim linearly decreases the learning rate, as we did not use this technique, the decay of the learning rate could help explain the noted differences. \\ The first hypothesis, the fact that we used a batched approach, may be confirmed by the fact that when combined with input shuffling SGD does perform closer to Gensim, going from 11 to 7 epochs to converge, as input shuffling reduces the number of co-occurrence of the same word in a batch.
Now the question arises if the 3 epochs, that Gensim is better, can be explained by the selection of better negative samples and the learning rate decay?

@@

\subsection{Future Work}
This work shows that the convergence time of the SGNS could be improved by using input shuffling and advanced optimizers. As with every work, there still exists possible extensions. First and foremost an aspect of our implementation that can be prejudicial is that we only extensively tested our model with one small dataset. The fact that we only used one dataset as well that it's a small one is problematic. \\ 
~~\textit{Problem with a small dataset:} \\ By using a very small dataset we do not use the model in the condition it is most needed for, as the dataset used in practice usually consists of more than 1 billion words. There is a small argument that can be made for machine translation as the use of small parallel corpus is not unusual in this field. \\
~~\textit{Problem with using only one dataset:}\\
It has been shown that some optimizers perform better with specific shapes of loss functions. To make a compelling argument it's, necessary to show that our model with the use of Input shuffling and Adam as its optimizer also outperform Gensim with other data sets, so that the claim can hold consistently.\\
Finally, our implementation did not outperform Gensim in run time, as this was not the goal of our work. Therefore, one could improve an already existing, optimized version, with input shuffling and advanced optimizers and should achieve a better run time than Gensim.

>>>>

\subsubsection{Large differences with NAG and SGD when using shuffling}
SGD and NAG both have very different values when using shuffling in comparison to unshuffled input, as shown in figures \ref{fig:results_sgd} and \ref{fig:results_nag}. We do not only attribute those results to input shuffling but partially also to a good random initialization guess. Due to a lack of time these experiments were not replicate more than once.

\subsection{Comparison to Gensim}
This subsection will compare our finding extensively to Gensim \cite{gensim}. As explained in Section \ref{ssec:gensim}, Gensim is optimized to have a very high throughput, this made it possible to achieve a lot of computations. Furthermore, Gensim provides access to the loss and the resulting word embeddings, which facilitated the comparison process.
This subsection is structured as follows: first it will describe the used configuration of Gensim, second it will compare Gensim to our implementation of the SGNS with the use of SGD and finally compare our model with Adam to Gensim. A graphic showing the comparison of our models to Gensim can be found in Figure \ref{fig:gensim_vs_adam}.

\subsubsection{Configuration of Gensim}
To compare our-self in a correct  manner we used the same dataset, with the same preprocessing parameters, i.e subsampling and min count. The hyper parameters of the network (window size, embedding size, number of negative sample, exponent to which the unigram distribution, which decides how a negative sample is used, is raised) are equivalent to our parameters described in Section \ref{ssec:config}. The only difference with our model is the learning rate. Gensim has a starting learning rate of $0.025$ and linearly decrease it at every epoch to $0.0001$.

\subsubsection{Gensim vs. SGD}
As stated earlier, we are not going to compare this work to Gensim in run time. Gensim is heavily optimized and written in cython\footnote{https://rare-technologies.com/word2vec-in-python-part-two-optimizing/}, which is 23x faster than plain Numpy. Since we used PyTorch the difference is not that big, but still remains. As shown in Figure \ref{fig:gensim_vs_adam} the convergence time was not the same between our implementation and Gensim. There are different possible reasons why this could be the case:\\ First, our batched approach could hinder performance in terms of convergence time since our loss function is not exactly the same. We compute the loss for multiple training samples by taking the sum over each score which is individually used as a loss by Gensim.\\ Secondly, a difference to our implementation is the fact that Gensim checks whether negative samples are not equal to the context word. If that is the case Gensim selects a new random sample. Therefore, the learning of the input and output context is optimized. \\Finally, another possibility is the decay of the learning rate used by Gensim. In fact, decaying the learning rate has been proven in a lot of work to decrease the convergence time. Gensim linearly decreases the learning rate, as we did not use this technique, the decay of the learning rate could help explain the noted differences. \\ The first hypothesis, the fact that we used a batched approach, may be confirmed by the fact that when combined with input shuffling SGD does perform closer to Gensim, going from 11 to 7 epochs to converge, as input shuffling reduces the number of co-occurrence of the same word in a batch.
Know the question arises if the 3 epochs, that Gensim is better, can be explained by the selection of better negative samples and the learning rate decay.

@@

\subsection{Future Work}
This work shows that the convergence time of the SGNS could be improved by using input shuffling and advanced optimizers. As with every work, there still exists possible extensions. First and foremost an aspect of our implementation that can be prejudicial is that we only extensively tested our model with one small dataset. The fact that we only used one dataset as well that it's a small one is problematic. \\ 
~~\textit{Problem with a small dataset:} \\ By using a very small dataset we do not use the model in the condition it is most needed for, as the dataset used in practice usually consists of more than 1 billion words. There is a small argument that can be made for machine translation as the use of small parallel corpus is not unusual in this field. \\
~~\textit{Problem with using only one dataset:}\\
It has been shown that some optimizers perform better with specific shapes of loss functions. To make a compelling argument it's necessary to show that our model with the use of input shuffling and Adam as its optimizer also outperforms Gensim with other data sets.\\
Finally, as it wasn't the goal of our implementation to outperform Gensim in run time, one could improve an already existing, optimized version, with input shuffling and advanced optimizers and should achieve a better run time than Gensim.
